Wide Post-Mass Transfer White-Dwarf Binaries

### 2024-07-11 summaries/paper-draft-plain.tex
finish organizing current summary

\begin{tabular}{|cl|cl|}
  \hline
  \verb|kstar| & \textbf{evolutionary state} & \verb|evol_type| & \textbf{evolutionary change} \\
  \hline
  0 & MS $<0.7 \Msun$ & 1 & initial state \\
  1 & MS $>0.7 \Msun$ & 2 & kstar change \\
  2 & Hertzprung Gap & 3 & begin Roche lobe overflow\\
  3 & First Giant Branch & 4 & end Roche lobe overflow\\
  4 & Core Helium Burning & 5 & contact \\
  5 & Early AGB & 6 & coalescence \\
  6 & TP-AGB & 7 & begin common envelope \\
  7 & Naked Helium Star MS & 8 & end common envelope \\
  8 & Naked Helium Star Hertzprung Gap & 9 & no remnant leftover \\
  9 & Naked Helium Star Giant Branch & 10 & max evolution time \\
  10 & Helium White Dwarf & 11 & binary disruption \\
  11 & Carbon/Oxygen White Dwarf & 12 & begin symbiotic phase \\
  12 & Oxygen/Neon White Dwarf & 13 & end symbiotic phase \\
  13 & Neutron Star & 14 & blue straggler \\
  14 & Black Hole & 15 & supernova of primary \\
  15 & Massless Remnant & 16 & supernova of secondary \\
  \hline
\end{tabular}

@@

In our simulation, we investigate how accretion efficiency $\beta$ and initial separation $a_i$ affects the final separation $a_f$. We run a grid of three different $\beta$ — $\beta = 0.0$, $0.5$ and $1.0$, and a linear grid of $a_i$ from $2 \au$ to $8 \au$ with $400$ steps. In total, we simulate $1200$ binaries with different accretion limit and initial separation. For each system, after the simulation is finished, we select the first timestep, if any, that shows a WD+MS binary. We record the separation at this moment as the final separation $a_f$, and we compare them with the observed minimum separation $0.15 \au$ for post-mass transfer WD binaries.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.75\linewidth]{fig/stable_hi_log.png}
  \caption{COSMIC results of WD+MS binaries after stable mass transfer for $7\Msun + 1\Msun$ systems, with accretion efficiency $\beta = 0$, $\beta = 0.5$, and $\beta = 1$ respectively. Left panel: dependence of final separation $a_f$ of the WD+MS binary on initial separation $a_i$. Right panel: The corresponding companion mass in solar units in the WD+MS binary. The gap in both figure represents that no WD+MS system forms for that initial separation.}
  \label{stable_hi}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig/sample-distribution.png}
  \caption{The distribution of primary mass, companion mass, and orbital period in the sampled population using independent sampler. Left panel: Distribution of primary mass and companion mass in log scale. Right panel: Distribution of orbital period $\Porb$ in log scale.}
  \label{sample-distribution}
\end{figure}

\bibliography{reference.bib}
\bibliographystyle{apalike}

>>>>

\begin{tabular}{|cl|cl|}
  \hline
  \verb|kstar| & \textbf{evolutionary state} & \verb|evol_type| & \textbf{evolutionary change} \\
  \hline
  0 & MS $<0.7 \Msun$ & 1 & initial state \\
  1 & MS $>0.7 \Msun$ & 2 & kstar change \\
  2 & Hertzsprung Gap & 3 & begin Roche lobe overflow\\
  3 & First Giant Branch & 4 & end Roche lobe overflow\\
  4 & Core Helium Burning & 5 & contact \\
  5 & Early AGB & 6 & coalescence \\
  6 & TP-AGB & 7 & begin common envelope \\
  7 & Naked Helium Star MS & 8 & end common envelope \\
  8 & Naked Helium Star Hertzsprung Gap & 9 & no remnant leftover \\
  9 & Naked Helium Star Giant Branch & 10 & max evolution time \\
  10 & Helium White Dwarf & 11 & binary disruption \\
  11 & Carbon/Oxygen White Dwarf & 12 & begin symbiotic phase \\
  12 & Oxygen/Neon White Dwarf & 13 & end symbiotic phase \\
  13 & Neutron Star & 14 & blue straggler \\
  14 & Black Hole & 15 & supernova of primary \\
  15 & Massless Remnant & 16 & supernova of secondary \\
  \hline
\end{tabular}

@@

In our simulation, we investigate how accretion efficiency $\beta$ and initial separation $a_i$ affects the final separation $a_f$. We run a grid of three different accretion efficiency — $\beta = 0.0$, $0.5$ and $1.0$, and a linear grid of $a_i$ from $2 \au$ to $8 \au$ with $400$ steps. In total, we simulate $1200$ binaries with different accretion limit and initial separation. For each system, after the simulation is finished, we select the first timestep, if any, that shows a WD+MS binary. We record the separation at this moment as the final separation $a_f$, and we compare them with the observed minimum separation $0.15 \au$ for post-mass transfer WD binaries.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.75\linewidth]{fig/stable_hi_log.png}
  \caption{COSMIC results of WD+MS binaries after stable mass transfer for $7\Msun + 1\Msun$ systems, with accretion efficiency $\beta = 0$, $\beta = 0.5$, and $\beta = 1$ respectively. Left panel: dependence of final separation $a_f$ of the WD+MS binary on initial separation $a_i$. Right panel: the corresponding companion mass in solar units in the WD+MS binary. The gap in both figure represents that no WD+MS system forms for that initial separation.}
  \label{stable_hi}
\end{figure}

@@

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig/sample-distribution.png}
  \caption{The distribution of primary mass, companion mass, and orbital period in the sampled population using independent sampler. Left panel: distribution of primary mass and companion mass in log scale. Right panel: distribution of orbital period $\Porb$ in log scale.}
  \label{sample-distribution}
\end{figure}

To investigate how the final white dwarf mass and orbital period depends on \verb|lambdaf|, we simulation the same sampled population with five different $\lambda$ values, $\lambda = 5$, $15$, $25$, $35$, $45$. After that, we select the first time-step that satisfies \verb|evol_type = 8|, \verb|kstar_1| is WD, and \verb|kstar_2| is MS. We record the white dwarf mass $\MWD$ and the final orbital period $P_{\mathrm{orb}, f}$ from this moment. We also select the moment when the first Roche lobe overflow starts, and record the initial orbital separation $P_{\mathrm{orb}, i}$ and \verb|kstar_1| type at this moment. These variables will later help us study how the orbit widens, with respect to different parameter $\lambda$ and initial separation $a_i$.

\subsection{Results}
\begin{figure}
  \centering
  \includegraphics[width=0.6\linewidth]{fig/Mwd-P scatter.png}
  \caption{Scatter plot of white dwarf mass $\MWD$ and final orbital period after common envelope in log scale $\log_{10}\Porb$. Different colors indicates results for different $\lambda$ values. For each different $\lambda$ value the same initial population is used.}
  \label{Mwd-P-scatter}
\end{figure}

To show how the change in $\lambda$ affects the final orbital period, we plot final orbital period $\Porb$ against white dwarf mass $\MWD$ in Figure \ref{Mwd-P-scatter}. From Figure \ref{Mwd-P-scatter}, we note that the final orbital period is larger in general for larger $\lambda$. This can be clearly seen from the figure for $\MWD$ ranges in $0.2 \sim 0.7 \au$. This is also expected based on Equation \ref{ebind}, since for larger $\lambda$, less energy is required to unbind the common envelope, leading to more energy after mass transfer and a larger orbital period.

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig/kstar-map-rearr.png}
  \caption{Each panel: for this $\lambda$ value in the title, scatter plot of white dwarf mass $\MWD$ and final orbital period after common envelope in log scale $\log_{10}\Porb$. Different colors represent different kstar type of the primary at the start of mass transfer.}
  \label{kstar-map}
\end{figure}

To also show how the final separation $\Porb$ depends on \verb|kstar_1| type at the start of first RLOF, we plot $\Porb$ against $\MWD$ for different $\lambda$ separately, and use different color to distinguish between different \verb|kstar_1| type at the start of mass transfer. The plot is shown in Figure \ref{kstar-map}. In general, if the star is at later stages of evolution (larger \verb|kstar|) when mass transfer starts, then the resulting white dwarf mass $\MWD$ and $\Porb$ are also both larger. Note also that few systems start the mass transfer at \verb|kstar_1 = 2| (Hertzsprung gap) and \verb|kstar_1 = 4| (Core Helium Burning). This is because stars crosses Hertzsprung in a short timescale, leaving little time for interaction. Also, stellar radius is small when going through Core Helium Burning, reducing the possibility of interactions that starts when \verb|kstar_1 = 4|.

\begin{figure}
  \centering
  \includegraphics[width=0.6\linewidth]{fig/dif-kstar map.png}
  \caption{Scatter plot of difference in orbital period $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ against the white dwarf mass $\MWD$. Different colors indicates different kstar type of the primary at the start of mass transfer.}
  \label{dif-kstar map}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=\linewidth]{fig/dif-kstar two plot.png}
  \caption{Left panel: scatter plot of initial orbital period at the start of RLOF $P_{\mathrm{orb}, i}$ against the final white dwarf mass $\MWD$. The color of the dots indicates how the final orbital period $P_{\mathrm{orb}, f}$ changes with respect to $\lambda=35$ and $\lambda=5$. That is, the color indicates the magnitude of $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$. Right panel: scatter plot of $\log_{10} \Delta \Porb$ against the initial orbital period in log scale $\log_{10} P_{\mathrm{orb}, i}$ at the start of RLOF. Different colors indicates different kstar type of the primary at the start of RLOF.}
  \label{dif-kstar-exp}
\end{figure}

To visualize the difference in final orbital period $\Porb$, and how they corresponds to different \verb|kstar_1| type at the start of RLOF, we create plot the difference in orbital period $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ against the white dwarf mass $\MWD$ in Figure \ref{dif-kstar map}. From Figure \ref{dif-kstar map}, we can see that for large \verb|kstar_1| type at the start of RLOF, $\Delta P = P_{\mathrm{orb}}^{\lambda = 35} - P_{\mathrm{orb}}^{\lambda = 5}$ is also larger. To explain this, we plot the how final white dwarf mass $\MWD$, difference in final orbital period $\Delta \Porb$, and orbital period before mass transfer $P_{\mathrm{orb}, i}$ related to each other in Figure \ref{dif-kstar-exp}. We see that the larger initial orbital period $P_{\mathrm{orb}, i}$, the more significant effect does $\lambda$ have on the final orbital period. This is not unexpected since during mass transfer, the separation changes as $\dot{a} / a = \text{const}$. Hence $\dot{\Porb} / \Porb = \text{const}$. Therefore, the larger the initial orbital period, the larger it changes, and thus the larger effect does $\lambda$ have.

\newpage
\bibliography{reference.bib}
\bibliographystyle{apalike}
